---
title: "google_cert_r"
author: "Wes W."
date: '2021-06-19'
output:
  html_document:
    df_print: paged
description: Code samples from the Google data analytics cert
---

```{r necessary packages}
install.packages('tidyverse')
install.packages('palmerpenguins')
```

```{r load packages}
library(tidyverse)
library(palmerpenguins)
```

The cert course focuses on using ggplot2 for creating visualizations on data. ggplot2 is apart of the 'tidyverse' set of packages, so it's included already tidyverse.

The datasets being used are 'palmerpenguins' which is included in 'tidyverse' and is often used for teaching the foundations of data cleaning/manipulation and plotting in R. Another dataset is 'diamonds' which is included in the 'ggplot2' pacakge.  'ggplot2' is included in 'tidyverse'.

Basic construct for plots is **data** then (+) **geom** then (+) **aesthetics** then (+) **facets**. The *+* allows for chaining layers for the plot.

ggplot2 has a great [cheat sheet](https://github.com/rstudio/cheatsheets/blob/master/data-visualization-2.1.pdf).

### Basic empty plot
```{r starting point for plot - renders nothing; just loads data}
ggplot(data=penguins)
```

### Add a layer for rendering data with a geom
```{r load data and add geom layer}
ggplot(data=penguins) + 
  geom_point(mapping=aes(x=bill_length_mm, y=bill_depth_mm))
```

Here, the *aes()* is for defining the aesthetic for the plot. *geom_point* is the geom adding to the plot.  

```{r add color to the plot - variable in different colors}
ggplot(data=penguins) + 
  geom_point(mapping=aes(x=flipper_length_mm, y=body_mass_g, color=species))
```

Here, ggplot will automatically colorize the plot just by adding the color attribute to *aes()*

```{r add a shape to plot - variable as a shape}
ggplot(data=penguins) + 
  geom_point(mapping=aes(x=flipper_length_mm, y=body_mass_g, shape=species))
```

```{r do more with aesthetics - add color, size, and shape for the data points}
ggplot(data=penguins) + 
  geom_point(mapping=aes(x=flipper_length_mm, y=body_mass_g, shape=species, color=species, size=species))
```

```{r use the alpha paramter to give gradient to the data points}
ggplot(data=penguins) + 
  geom_point(mapping=aes(x=flipper_length_mm, y=body_mass_g, alpha=species))
```

Colors can also be defined directly
```{r set the color directly}
ggplot(data=penguins) + 
  geom_point(mapping=aes(x=flipper_length_mm, y=body_mass_g, alpha=species), color="orange")
```

Geoms can be combined because they are simply layered with each other onto the plot
```{r combined plot with two geoms}
ggplot(data=penguins) + 
  geom_smooth(mapping=aes(x=flipper_length_mm, y=body_mass_g)) +
  geom_point(mapping=aes(x=flipper_length_mm, y=body_mass_g))
```

Bar chart is a common graph.  This is how it's coded in R.
```{r plot with bar chart geom}
ggplot(data=diamonds) + 
  geom_bar(mapping=aes(x=cut, fill=cut))
```

Setting *fill* to a data point will enable geom_bar to select colors automatically for each data point.

```{r plot with bar chart geom different data point}
ggplot(data=diamonds) + 
  geom_bar(mapping=aes(x=cut, fill=clarity))
```

## Facets
Facets are subsets or sub-groupings of the entire dataset. By specifying the data point to be a facet, ggplot2 will automatically split/sub-divide the dataset. A facet is added just like any other layer, with the *+* character.

There are two types of facets: *facet_wrap()* and *facet_grid()*.

Using *facet_wrap()*
```{r adding a facet to penguins}
ggplot(data=penguins) + 
  geom_point(mapping=aes(x=flipper_length_mm, y=body_mass_g, color=species)) +
  facet_wrap(~species)
```

```{r adding a facet to diamonds}
ggplot(data=diamonds) + 
  geom_bar(mapping=aes(x=color,fill=cut)) +
  facet_wrap(~cut)
```

Using *facet_grid()*
```{r adding facet to penguins }
ggplot(data=penguins) + 
  geom_point(mapping=aes(x=flipper_length_mm,y=body_mass_g,color=species)) +
  facet_grid(species~sex)
```

## Annotations
Adding annotations will create another layer with extra information such as title, callouts etc

Adding title and citation information can be added with *labs()*.

```{r plot with labeling}
ggplot(data=penguins) + 
  geom_point(mapping=aes(x=flipper_length_mm, y=body_mass_g, color=species)) +
  labs(title="Palmer Penguins: Body Mass vs. Flipper Length", 
       subtitle="Sample of 3 peaguin species",
       caption="Data collected by Dr. Kristen Gorman")
```

*annotate()* can be used for adding labeling/information inside of the plot itself.

```{r plot with annotation}
ggplot(data=penguins) + 
  geom_point(mapping=aes(x=flipper_length_mm, y=body_mass_g, color=species)) +
  labs(title="Palmer Penguins: Body Mass vs. Flipper Length", 
       subtitle="Sample of 3 peaguin species",
       caption="Data collected by Dr. Kristen Gorman") +
  annotate("text", x=220, y=3500, label="The Gentoos are the largest", 
           color="purple", fontface="bold", size=4.5, angle=25)
```

## Finally
Of course, all of these plots, data, and layers can be assigned to variables  for reuse

```{r}
penguin_plot <- ggplot(data=penguins)
penguin_geom <- geom_point(mapping=aes(x=flipper_length_mm, y=body_mass_g, color=species))
penguin_labs <- labs(title="Palmer Penguins: Body Mass vs. Flipper Length", 
       subtitle="Sample of 3 peaguin species",
       caption="Data collected by Dr. Kristen Gorman")
penguin_annot <- annotate("text", x=220, y=3500, 
                        label="The Gentoos are the largest", color="purple", 
                        fontface="bold", size=4.5, angle=25)

penguin_plot + penguin_geom + penguin_labs + penguin_annot

```

